---
title: "Step 3 Project 1"
format: html
editor: visual
---

## Step 3 Project 1

The World Bank’s SDG database is a subset of its broader World Development Indicators, organized specifically around the UN’s 17 Sustainable Development Goals. It provides country-year panel data (typically yearly) covering economic, social, environmental, and institutional statistics aligned with each goal. The SDGs are separated by topic (e.g., poverty, education, gender equality). 

The proportion of women in national parliaments variable is under the 5th SDG. The 5th SDG is "Achieve gender equality and empower all women and girls." This variable measures the percentage of seats held by women in a country's national parliament, typically focusing on the lower or single legislative chamber. It is reported annually and serves as an indicator of gender equality in politics. 

This proportion is available for most countries from 1997-2020. There are missing values for certain years and countries, but the proportion is more available during the later years of that range.

\- \[SDG 5 description\](https://sdgs.un.org/goals/goal5)

\- \[Where we found the data\](https://databank.worldbank.org/source/sustainable-development-goals-(sdgs)#)

\- \[Variable description from World Bank\](https://databank.worldbank.org/metadataglossary/world-development-indicators/series/SG.GEN.PARL.ZS)

``` r
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(glue)
library(readxl)
```

```{r}
tuesdata <- tidytuesdayR::tt_load('2024-11-05')
democracy_data <- tuesdata$democracy_data
```

## GDP Data

```{r}
library(httr)
library(tidyjson)
library(jsonlite)
library(tidyverse)
library(glue)
library(readxl)
library(countrycode)
```

```{r}
democracy_data <- readr::read_csv('https://raw.githubusercontent.com/rfordatascience/tidytuesday/main/data/2024/2024-11-05/democracy_data.csv')
```

```{r}
#| eval: True
my_key <-"jm6LTqDYOeb7igZXbOxfolhV4eNVaznj2yQqj2EC"
```

```{r}
get_gdp <- function(country, key = my_key)
  {
  url <- glue("https://api.api-ninjas.com/v1/gdp?country={country};X-Api-Key={key}")
  my_data <- safely(fromJSON)(url)
  
  parsed <- my_data$result
  
  if (is.null(parsed) || length(parsed$year) == 0) {
    return(NULL)
  } else {
    return(
      data.frame(
        country = country,
        year = my_data$result$year,
        gdp_nominal = my_data$result$gdp_nominal
      )
    )
  }
  }
```

```{r}
#Data cleaning: Renaming country_code column to just country to make it easier to join the datasets
democracy_data <- democracy_data %>% 
  rename(country = country_code) 
```

```{r}
gdp_data <- democracy_data |> 
  distinct(country) |>
  pull(country) |> 
  map(get_gdp) |> 
  bind_rows() |>
  filter(year <= 2025)
```

```{r}
democracy_data <- democracy_data |>
  left_join(gdp_data, by = c("country", "year"))
```

## Data Cleaning

Below, we read in the data and do some data cleaning. We removed some columns because certain years had no data for any countries, and there was an empty column that was created from downloading the data. Lastly, we renamed the country column.

```{r}
womenPar0 <- read_excel("womenPar.xlsx", skip=2)
womenPar0 <- womenPar0 %>%
  select(-2:-8, -11) %>% # removing columns with no data and empty column
  rename(country = ...1) # renaming the country column
womenPar0
```

This data frame is in a wide format, so we used pivot_longer() to pivot the data into long format to match our democracy_data.

```{r}
womenPar <- womenPar0 %>%
  pivot_longer(cols = -country, # keep country column the same
    names_to = "year", # creating new column for years
    values_to = "prop_in_par") # creating new column for the proportion data
womenPar <- as.data.frame(womenPar)
womenPar
```

# Joining the Dataset to the Democracy Data

```{r}

#changing the data from the womenPar data to have years and proprotion in parliament be numeric, and change the na values to be repalced with na instead of "..". 
#the package countrycode can take in a vector, such as a column from our data frame in womenPar, and convert it to an iso3c, which is a country code. By adding that column, we are able to achieve a join much easier.
womenPar <- womenPar |>
  mutate(
    year = as.numeric(year),
    prop_in_par = na_if(prop_in_par, ".."),
    prop_in_par = as.numeric(prop_in_par),
    iso3c = countrycode(country, "country.name", "iso3c")
  )
#making the year numeric too, and then changing the countrycode column to be called iso3c to match our womenPar data.
democracy_data <- democracy_data |>
  mutate(
    year = as.numeric(year),
    iso3c = countrycode(country_name, "country.name", "iso3c")
  )
#using a left join to join all our data sets together, since the GDP one is already included. We are joining by the code and the year, since each year is what is every row, and the code is the country recognition.  
democracy_data <- democracy_data |>
  left_join(womenPar |> select(iso3c, year, prop_in_par),
            by = c("iso3c", "year"))
```
